Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-09-09 draft.tex
Revised part to new IVA tool

\subsection{Interactive Display of Decision Trees}
We have to create a decision tree for every non-image variable to analyze which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields a corresponding number of trees.
%%
Dividing the subjects into four groups using body size, the number increases to 402 trees.
%%
%To make the results cognitive feasible, we have to abstract the results of the classification.
We have to abstract the results of the classification to keep the mental effort of interpreting the data low.
%%------------
\subsubsection{Visualization of Classification Results}
\label{subsec:VisualizationOfClassificationResults}
%%
%We follow the visual information-seeking mantra and want to provide an overview first, then details-on-demand \cite{shneiderman1996}.
We follow the Visual Analytics mantra by analyzing first, show the important and analyze further \cite{Keim}.
%%
A first analysis step was performed by applying the classification algorithm to the data.
%%
The optimal classification uses a few rules to precisely characterize the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error rate}.
%%
The two measures form the axis for a \emph{scatter plot} of the classification results.

@@

Many follow-up variables are sparse, such as \emph{medication of diabetes} or \emph{reason of early retirement}.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size.
%%
This makes these results less reliable.
%%
We removed variables with less than 100 subjects, since they have too few subjects to be statistically significant.
%%
Furthermore, we map the number of subjects associated with a variable to point diameter in the scatter plot.
%%
This allows to instantly assess the reliability of the result.
%%

@@

%\com{ToDo: SQRT Skalierung der Treesize-Dimension, da hauptsaechlich geringe Groessen von interesse sind, ausreisser aber diesen Bereich schwer durchschaubar machen}
% %%
%
% %%
% Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatter plot.
% %%
% By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).

@@

\subsubsection{Dummy Variables}
Dummy variables convert a categorical variable with multiple manifestations into several binary variables.

@@

Each binary variables encodes the presence of a manifestation.

@@

For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four binary variables (\emph{No pain - Yes/No} to \emph{Large Pain - Yes/no}).

@@

One subject can only have one of these variables set to true.
This is useful for our classification, because it allows to determine which manifestation of a variable can be described best using the image data parameter.
%%\com{ToDo: Dummy-Variables}
\subsubsection{Interaction With the Visualization}
The described visualization provides a good overview over the classification results.
We still want to be able to display \emph{details-on-demand} \cite{shneiderman1996} and examine a decision tree in detail.
This is realized by clicking on an entry on the visualization, which then displays the corresponding decision tree in detail.
This allows for sequentially analyzing the classifications.

@@

The interactive visualizations are realized using the \texttt{ggvis}\footnote{GGVIS Footnote} package.

@@

We used \texttt{R Shiny}\footnote{R Shiny footnote} to make the report available as dynamic web application.

>>>>

\subsection{Interactive Display of Decision Trees}
We have to create a decision tree for every non-image variable to analyze which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields a corresponding number of trees.
%%
Further subdivision, e.g. by quantiles of \emph{body size} increases the number to 402 trees.
%Dividing the subjects into four groups using body size, the number increases to 402 trees.
%%
%To make the results cognitive feasible, we have to abstract the results of the classification.
We have to abstract the results of the classification to keep the mental effort of interpreting the data low.
%%------------
\subsubsection{Visualization of Classification Results}
\label{subsec:VisualizationOfClassificationResults}
%%
%We follow the visual information-seeking mantra and want to provide an overview first, then details-on-demand \cite{shneiderman1996}.
We follow the Visual Analytics mantra by analyzing first, show the important and analyze further \cite{Keim}.
%%
A first analysis step was performed by applying the classification algorithm to the data.
%%
The optimal classification uses a few rules to precisely characterize the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error rate}.
%%
The two measures form the axis for a \emph{scatter plot} of the classification results.

@@

Many follow-up variables are sparse, such as \emph{medication of diabetes} or \emph{reason of early retirement}.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size.
%%
This makes these results less reliable.
%%
%We removed variables with less than 100 subjects, since they have too few subjects to be statistically significant.
Therefore we provide a way to adjust the minimal number of subjects for each variable using a slider input.
%%
The initial value is empirically set to $100$, marking a good tradeoff between sparse variables and statistical informative value.
%%
Furthermore, we map the number of subjects associated with a variable to point diameter in the scatter plot.
%%
This allows to instantly assess the reliability of the result.
%%

@@

%\com{ToDo: SQRT Skalierung der Treesize-Dimension, da hauptsaechlich geringe Groessen von interesse sind, ausreisser aber diesen Bereich schwer durchschaubar machen}
% %%
%
% %%
% Abstracting the classification results into two continuos variables allows us to visualize the result of 134 classifications in one scatter plot.
% %%
% By mapping the group affiliation (in our case body-size) to color or shape, we can even plot all 402 results in a single visualization (Fig.~\ref{fig:decision-tree} a).
%%
\subsubsection{Dummy Variables}
Dummy variables convert a categorical variable with multiple manifestations into several binary variables.
%%
Each binary variables encodes the presence of a manifestation.
%%
For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four binary variables (\emph{No pain - Yes/No} to \emph{Large Pain - Yes/no}).
%%
One subject can only have one of these variables set to true.
%%
This is useful for our classification, because it allows to determine which manifestation of a variable can be described best using the image data parameter.
%%\com{ToDo: Dummy-Variables}
%%------------
\subsubsection{Interaction With the Visualization}
%%
The described visualization provides a good overview over the classification results.
%%
We still want to be able to display \emph{details-on-demand} \cite{shneiderman1996} and examine a decision tree in detail.
%%
This is realized by clicking on an entry on the visualization, which then displays the corresponding decision tree in detail.
%%
This allows for sequentially analyzing the classifications.

We also provide controls for adjusting the maximum classification error and minimum subject count for a variable.

@@

This gives the user control to abstract or refine the displayed information.

@@

Selecting a variable using a drop down menu allows to user to select the variable used for subject subdivisions, e.g. \emph{gender} or \emph{disease treatment type}.

@@

Categorization using metric variables, such as \emph{body size}, is done using their quantiles.

@@

This allows to assess the influences of a variable to the classification process.

@@

The interactive visualizations are realized using the \texttt{ggvis}\footnote{Developed by RStudio, Inc; \texttt{ggvis.rstudio.com}} package.

@@

We used \texttt{R Shiny}\footnote{Developed by RStudio, Inc; \texttt{shiny.rstudio.com}} to make the report available as dynamic web application.
